---
title: 'Regression Week 4 Project: Motor Trends : Automatic or Manual transmission?'
output:
  pdf_document: default
  html_document:
    df_print: paged
---



You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
"Quantify the MPG difference between automatic and manual transmissions"  


Peer Grading

The criteria that your classmates will use to evaluate and grade your work are shown below.
Each criteria is binary: (1 point = criteria met acceptably; 0 points = criteria not met acceptably)
Criteria

Did the student interpret the coefficients correctly?
Did the student do some exploratory data analyses?
Did the student fit multiple models and detail their strategy for model selection?
Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
Did the student do a residual plot and some diagnostics?
Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?
Did the report include an executive summary?
Was the report done in Rmd (knitr)? 


### Exploring the Dataset

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
data(mtcars)
```



```{r, warning=FALSE}
# Checking for the dimension of the dataset
dim(mtcars)
```

```{r, warning=FALSE}
# Checking for columns names
names(mtcars)
```



```{r}
# viewing the first six rows of the dataset
head(mtcars)
```


```{r}
# checking the structure of the dataset
str(mtcars)
```

```{r}
# checking the summary of the dataset
summary(mtcars)
```


```{r}
# checking the correlation between mpg and the other variable
cor(mtcars$mpg, mtcars[,-1])
```

The result of the correlation provides a negative correlation to mpg and the variables are:
*cyl, disp, hp, wt, carb*


```{r}
# Looking at the t.test at the confident level of 95 and the p value
mtcarsTestResults <- t.test(mpg ~ am, data = mtcars, conf.level = 0.95)
mtcarsTestResults$p.value
```

If the null hypothesis is presented as that a car with a manual transmission will have more mpg than a car with an automatic transmission. However, the p value been 0.001374 may reject the null hypothesis. More over, as the cars do have different characteristics a deeper analysis must be pursue.

```{r}
# Checking the estimate of t.test results
mtcarsTestResults$estimate
```

The result from the estimate shows that the manual transmission has an advantage of 7.24494 extra miles per gallon  


#### Regression Analysis

```{r}
# Checking the regression values of the univariable mtcarsfitUv
mtcarsfitUv <- lm(mpg ~ am, data = mtcars)
summary(mtcarsfitUv)
```

```{r}
# Rounding the univariable mtcarsfitUv to two digits
round(summary(mtcarsfitUv)$r.square*100, digits=2)
```

The result of the regression can only explain that 35.98% of the mpg variation

```{r, results='hide'}
# Hidden results to comply with the page constraint.}
# Checking the regression values of the multivariable mtcarsfitMv
mtcarsfitMv <- lm(mpg ~ ., data = mtcars)
mtcarsfitMv
summary(mtcarsfitMv)
```

 
```{r, results='hide'}
# Hidden results to comply with the page constraint.
stepFit <- step(mtcarsfitMv)
summary(stepFit)
summary(stepFit)$coeff

```


```{r}
# Rounding the multivariable mtcarsfitMv to two digits
round(summary(mtcarsfitMv)$r.square*100, digits=2)
```

#### Summary

We use the mtcars dataset to determine which type of transmission would offer better mileage.
The t-test between the automatic and manual transmission shows us that the maul transmission does have a 7.24494 extra mileage per gallon.
However, the regression model on a multivariable shows that the difference between automatic and manual transmission is much less 

#### Conclusion

Although the manual transmission would have a small advantage than an automatic transmission; there are other variable like cyl, wt, hp which have more more influence when determine

#### Residuals Analysis and Diagnostics

- The residual vs fitted plots shows not consistent pattern
- The Normal Q-Q plot indicates that the residuals are normal
- The scale-location plot confirms the constant variance assumption
- The Residuals vs leverage confirms that no outliers are present

### Appendix

#### 1. 

```{r}
# Creating a new variable as a factor and changing the levels
mtcars$am2 <- factor(mtcars$am)
levels(mtcars$am2) <-c("Automatic", "Manual")

# Using ggplot2 to create a boxplot to look at mpg and transmission type distribution
g <- ggplot(mtcars,aes(x=am2, y=mpg)) +
  geom_boxplot(aes(group=am2), fill=c("lightblue","salmon")) +
  labs(x="Transmission Type") +
  ylab("Miles per Gallon") +
  ggtitle("Miles per Gallon vs Transmission Type") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_jitter(cex = 1.8, shape = 5, aes(colour = mpg))
g
```

#### 2. 

```{r}
# pairs all the variables without the newly create variable am2
pairs((mtcars %>% select(-am2)), panel = panel.smooth, main = "Pair Graph of mtcars")
```

The Pair graph shows that several variables are high correlated with mpg.

#### 3.  


```{r}
# Multivariable
par(mfrow=c(2, 2))
plot(mtcarsfitMv)
```

